Modeling browser-based distributed evolutionary computation systems

### 2015-02-08 modeling-volunteer-computing-gecco.tex
New data

>>>>

\subsection{Server side}
\label{ss:server}

The server was also written using node.js and the {\tt express.js}
module and has three REST \cite{DBLP:journals/corr/abs-1105-4971}
functions, that is, three functions that can be accessed from any
client, including the browser, via a request encoded in an URL and a
standard HTTP command. These functions are\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\item {\tt GET log} returns the experiment log so far. It is used by
  the researcher to gather data, and as a transparency measure so that
  anyone can use it. 
\end{itemize}

In fact, the program includes more logic to serve the static pages
that contain the talk in which it is included and the JavaScript/CSS
code also, but it all amounts to a small amount of code. This code can
be deployed, in principle, in any server running node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a freeriding way, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014}

Having the pages and the server in the same domain allows to work in
default Ajax mode, that does not allow cross-site requests. However,
these can also be disabled if needed. In this case, every domain holds
a single experiment, there is no multi-tenancy of experiments, so its
management is quite simple.

When the experiment is finished, that is, the solution has been found,
the server is restarted using  command line tools provided by
Heroku. Every experiment was first done during the talk and then
announced in Twitter several times until enough data was gathered.

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in a GitHub
repository (address hidden for double blind review).
